Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-28 manuscript/manuscript.Rmd
appendices A and B

Finally, in our study case, we have shown that the ‘‘rest periods’’ so prominent in both the DRL solution and in the optimal bang-bang policy -- periods of several time steps with no fishing at all -- may be important to maintain the sustainability of the fishery and to increase its long-term profits.
This hypothesis could be explored in more generality as it might apply to a wide variety of models with different optimal control strategies.
In App. \ref{app:tbd} we provide some elementary support for this hypothesis by examining the performance of several ‘‘rule of thumb’’ policies based on conservative corrections to MSY.

>>>>

Finally, in our study case, we have shown that the ‘‘rest periods’’ so prominent in both the DRL solution and in the optimal bang-bang policy -- periods of several time steps with no fishing at all -- may be important to maintain the sustainability of the fishery and to increase its long-term profits.
This hypothesis could be explored in more generality as it might apply to a wide variety of models with different optimal control strategies.

# Appendix A: collected time series data

In this appendix we provide a wider selection of time series taken from the same data set as Fig. \ref{fig:timeseries}.
Namely, Figs. \ref{fig:timeseries_msy}, \ref{fig:timeseries_tac} and \ref{fig:timeseries_ppo} display these time series.

```{r} 
#| label: timeseries_msy
#| fig.cap: "Nine time series of the system controlled by the MSY constant mortality policy." 
#| out.width: "6.5in"
msy_tac_ppo |> 
    group_by(rep) |>
    filter(rep < 9, policy == "msy") |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(rep)) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 0.6)
```

```{r} 
#| label: timeseries_tac
#| fig.cap: "Nine time series of the system controlled by the policy using 80 percent of the MSY constant mortality." 
#| out.width: "6.5in"
msy_tac_ppo |> 
    group_by(rep) |>
    filter(rep < 9, policy == "tac") |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(rep)) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 0.6)
```

```{r} 
#| label: timeseries_ppo
#| fig.cap: "Nine time series of the system controlled by the trained DRL agent." 
#| out.width: "6.5in"
msy_tac_ppo |> 
    group_by(rep) |>
    filter(rep < 9, policy == "ppo") |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(rep)) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 0.6)
```

# Appendix B: Conservative scale-downs of MSY

Here we show the performance of constant mortality policies with a fraction $\alpha$ of the MSY mortality rate (in the main text we focused on $\alpha=0.8$).
Their performance is evaluated in Fig. \ref{fig:several-tac}.
Notice the sharp transition in performance between $\alpha = 0.9$ and $0.95$.
While the behaviour of the former closely mimics, e.g., the $\alpha=0.8$ policy, the performance of the latter is essentially the same as the optimal constant mortality ($\alpha=1$, see Fig. \ref{fig:rew_len_fig}).
This points at the fact that the value of the parameter $\alpha$, although commonly chosen simply as a rule of thumb, lives on a ‘‘knife's edge,’’ small changes in this parameter may lead to drastic changes in the dynamics of the controlled system.
What is worse: the particular scale at which this transition happens is generally dependent on the system's intrinsic dynamics.
In reality, where fishery models and their respective parameters are subject to substantial uncertainty, this fact calls to question the purported additional safety obtained via these rules of thumb.

```{r}
#| label: several-tac
#| fig.cap: "Performance of constant mortality policies with mortality $\\alpha M^*$ for several values of $\\alpha$.
#| The top panel displays the episode lengths for each repetition, while the bottom pannel displays the episode harvest.
#| The labels in the plots are the value of $\\alpha$ expressed as a percentage."
#| out.width: "6.5in"
N <- 30
actions <- unique(msy_df$action)
i85 <- which.min(abs(actions - best_action* 0.85))
i90 <- which.min(abs(actions - best_action* 0.90))
i95 <- which.min(abs(actions - best_action* 0.95))
tac_85_sim <- msy_df |> 
    filter(action == actions[[i85]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "85%")
tac_90_sim <- msy_df |> 
    filter(action == actions[[i90]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "90%")
tac_95_sim <- msy_df |> 
    filter(action == actions[[i95]]) |>
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1, policy = "95%")

tac_80_sim <- tac_sim |> mutate(policy = "80%")

tac4 <- bind_rows(tac_80_sim, tac_85_sim, tac_90_sim, tac_95_sim)

p1 <- tac4 |> 
    group_by(rep, policy) |>
    filter(t==max(t), species == "X", rep < N) |>
    ggplot(aes(rep, t)) + 
      geom_col() +
      facet_wrap(~factor(policy), ncol = 4) +
      labs(y = "episode length", x = "repetition") +
      theme(axis.text.x = element_blank())

p2 <- tac4 |> 
    group_by(rep, policy) |>
    filter(t==max(t), species == "X", rep < N) |>
    ggplot(aes(rep, reward)) + 
      geom_col() +
      facet_wrap(~factor(policy), ncol=4) +
      labs(y = "episode harvest", x = "repetition") +
      theme(axis.text.x = element_blank())

(p1 / p2)
```
